Mathematical Aspects of the Capital Market

### 2017-07-21 phd/Chapterintroduction.tex
Added my own thesis.

In this thesis we study the largest and smallest eigenvalues of high-dimensional sample covariance and correlation matrices of heavy-tailed time series.

\section{Random Matrix Theory}\label{sec:rmt}

The field of Random Matrix Theory (RMT) is concerned with the spectral properties of high-dimensional random matrices. Its development was motivated by applications.
In quantum mechanics, for example, the energy levels of particles in a large system can be characterized by the eigenvalues of a random infinite-dimensional Hermitean operator $\bfW$ on a Hilbert space. It is common to work with discretizations in a finite-dimensional space. In this case $\bfW$ turns into a high-dimensional Hermitean random matrix. Such matrices are called Wigner matrices named after Eugene Paul Wigner. He proved in the $1950$s that 
if $\bfW$ has independent standard normal entries on and above the diagonal, then
the expected empirical distribution of the eigenvalues of $\bfW$ tends to the so-called semi-circle law defined in \eqref{eq:semi-circle}, as the dimension of $\bfW$ goes to infinity; see for example \cite{wigner:1955,wigner:1957}. Since then a great variety of asymptotic results has been proved for various classes of random matrices under different assumptions on the distribution of the entries and their dependence structure. The discovery of many results was triggered by the enormous improvement of computation power which led to numerous conjectures. In the second half of the $20$th century, the research on asymptotic spectral properties of large-dimensional random matrices attracted considerable interest among physicists, computer scientists and mathematicians.  
A breakthrough in the theory of spectral distributions of sample covariance matrices was achieved by Mar\v cenko and Pastur in $1967$. 
%In $2010$, the most famous conjecture about the circular law was solved in \cite{tao:vu:krishnapur:2010}.

For many years the main focus of research in RMT has been on limiting spectral distributions. More recently, the focus turned to linear spectral statistics, eigenvectors, limiting distributions of extreme eigenvalues and their spacings. This thesis addresses these four topics in the special setting of random matrices with  heavy-tailed entries.

RMT is a versatile and useful tool in many fields of modern sciences that are faced with high-dimensional data sets. It  employs techniques from probability theory, multivariate statistics, number theory and combinatorics. Moreover, RMT finds applications in quantum physics, signal processing, wireless communications and finance; see Bai et al.~\cite{bai:fang:liang:2014} for more detailed examples. 
%``Big data'' is a popular topic nowadays. To meet the demand for data analysis and dimension reduction techniques RMT brings together

\subsection{Limiting spectral distributions}

For any random $p\times p$ matrix $\bfA$ with real eigenvalues $\lambda_1(\bfA),\ldots,\lambda_p(\bfA)$ the {\em empirical spectral distribution} is defined by 
\begin{equation*}
F_{\bfA}(x)= \frac{1}{p}\; \sum_{i=1}^p \1_{\{ \lambda_i(\bfA)\le x \}}, \qquad x\in  \R\,.
\end{equation*}

A major problem in random matrix theory is to find the weak limit of $(F_{\bfA_n})$, the so-called {\em limiting spectral distribution}, for suitable sequences of Hermitean $p\times p$ matrices $(\bfA_n)$. By weak convergence of a sequence of probability distributions $(F_{\bfA_n})$ 
to a \pro y \ds\ $F$, we mean $\lim_{\nto} F_{\bfA_n}(x)=F(x) \,\as$ for all continuity points of $F$.
Although the eigenvalues of $\bfA_n$ are continuous functions of the entries of $\bfA_n$ there are no closed-form expressions if the dimension is larger than $4$. Therefore methods to identify and characterize the limiting spectral distribution are needed. We will briefly discuss the two most common ones: the method of moments and Stieltjes transforms.
\par

By Lemma B.3 in \cite{bai:silverstein:2010}, a distribution function $F$ is uniquely characterized by its sequence of moments 
\begin{equation*}
\beta_k=\int_0^{\infty} x^k \dint F (x)\,, \quad k=1,2,\ldots
\end{equation*}
if Carleman's condition $\sum_{k=1}^{\infty} \beta_{2k}^{-1/(2k)}=\infty$ is satisfied. In this case, weak convergence of $(F_{\bfA_n})$ to $F$ is equivalent to the convergence of moments, that is
\begin{equation}\label{eq:mmom}
\beta_k(\bfA_n)= \int_0^{\infty} x^k \dint F_{\bfA_n} (x)=\frac{1}{p} \tr(\bfA_n^k) \to \beta_k\,,\quad \nto, \,k=1,2,\ldots
\end{equation}
In many cases the calculation of $\tr(\bfA_n^k)$ is very demanding. Often its mean and variance are estimated by combinatorial techniques.
On the positive side, if $F$ has finite support, Carleman's condition holds automatically.
\par

Another useful tool is the {\em Stieltjes transform}
of the empirical spectral distribution $F_{\bfA}$:
\begin{equation*}
s_{F_{\bfA}}(z)= \int_{\R} \frac{1}{x-z} \dint F_{\bfA}(x) = \frac{1}{p} \tr\big((\bfA -z \bfI)^{-1}\big)\,, \quad z\in \C^+\,,
\end{equation*}
where $\C^+$ denotes the complex numbers with positive imaginary part. 
Weak convergence of $(F_{\bfA_n})$ to a distribution function $F$ is equivalent to $s_{F_{\bfA_n}}(z) \to s_F(z)$ a.s. for all $z\in \C^+$.
Notice that the Stieltjes transform $s_F$ determines a distribution function $F$ at all continuity points $a,b$ of $F$ via
\begin{equation*}
F(b)-F(a)= \lim_{v\to 0^+} \frac{1}{\pi} \int_a^b \Im s_F(x+\mathrm{i} v) \dint x\,.
\end{equation*}

\section{Sample covariance matrices}
For a sample of $n$ column vectors $\x_1,\ldots,\x_n$ of a $p$-dimensional time series the (non-normalized) {\em sample covariance matrix} is usually defined as
\begin{equation*}
\bfS=\sum_{i=1}^n (\x_i-\overline{\x})(\x_i-\overline{\x})'= \sum_{i=1}^n \x_i \x_i' - n\overline{\x}\,\overline{\x}'= \X\X' - n\overline{\x}\,\overline{\x}'\,,
\end{equation*}  
where $\overline{\x}=n^{-1}\sum_i \x_i$ is the sample mean and 
\begin{equation*}
\X=(\x_1, \ldots, \x_n) =(X_{it})_{i=1,\ldots,p;\,t=1,\ldots,n}
\end{equation*}
is the {\em data matrix}. The sample covariance matrix is of crucial importance in multivariate statistics, for instance in principal component analysis, canonical correlation analysis, multivariate regression, factor analysis, hypothesis testing and discriminant analysis. The case of multivariate normal observations has played a particular role in the development of statistical theory. Already in $1928$, Wishart \cite{wishart:1928} studied sample covariance matrices with normal entries. Through the $20$th century non-asymptotic procedures for Gaussian observations such as Fisher's test, Student's test, and the analysis of variance were developed. In practice, however, observations are often not normally distributed and asymptotic methods based on limit theorems for certain model parameters are employed instead of exact results which are difficult to obtain.

Most of the classical limit theorems are derived under the assumption that the dimension $p$ is fixed and the sample size $n$ goes to infinity. 
If the assumptions of the law of large numbers are satisfied, then $n^{-1} \bfS$ converges \as~to the covariance matrix $\Sigma$ of $\x_1$.

If $p$ is moderately large, it is known that $n^{-1} \bfS$ ceases to be a good estimate for $\Sigma$. This means that classical methods based on fixed dimension and large sample limits may lead to wrong conclusions when applied to high-dimensional data. One would need appropriate adjustments. RMT provides limit theory in the case of large $p$. New statistical methods can be built on these results.

In RMT one assumes that $p=p_n$ grows with $n$. The most common condition in the literature is
\begin{equation}\label{Cp}
\lim_{\nto} \frac{p_n}{n}\to \gamma\in (0,\infty)\,. %\tag{$G_{\gamma}$}
\end{equation}
The asymptotic spectral behaviors of two large matrices are the same if their difference is of finite rank. Therefore we will refer to $\X\X'$ as the {\em sample covariance matrix} from now on. Indeed, by the rank inequality (see Bai and Silverstein \cite[Theorem A.44]{bai:silverstein:2010}) we have for the supremum norm
\begin{equation*}
\| F_{n^{-1}\bfS}-F_{n^{-1}\X\X'} \| \le p^{-1} \mathrm{rank}((\overline{\x},\ldots,\overline{\x}))\,,
\end{equation*}
which means that the respective limiting spectral distributions coincide under \eqref{Cp}. For the same reason we can assume without loss of generality that the entries $(X_{it})$ are centered whenever the expectations exist. 
\par

The limiting spectral distribution of normalized sample covariance matrices was found by Mar\v cenko and Pastur.

\begin{theorem}[Debashis and Aue \cite{debashis:aue:2014}]\label{thm:mpch1}
Suppose that $\X$ has iid entries with common mean and variance $1$. If \eqref{Cp} holds, then, with probability one, $(F_{n^{-1}\X \X'})$ converges weakly to a non-random distribution, the so-called  Mar\v cenko--Pastur law $F_\gamma$. If $\gamma \in (0,1]$,  $F_\gamma$  has density,
\begin{eqnarray}\label{eq:MPch1}
f_\gamma(x) =
\left\{\begin{array}{ll}
\frac{1}{2\pi x\gamma} \sqrt{(b-x)(x-a)} \,, & \mbox{if } a\le x \le b, \\
 0 \,, & \mbox{otherwise,}
\end{array}\right.
\end{eqnarray}\noindent
where $a=(1-\sqrt{\gamma})^2$ and $b=(1+\sqrt{\gamma})^2$. If $\gamma>1$, the \MP law is a mixture of a point mass at $0$ and the density function $f_{1/\gamma}$ with weights $1-1/\gamma$ and $1/\gamma$, respectively.
\end{theorem}

The \MP law describes the global behavior of the eigenvalues of $\X \X'$. Theorem \ref{thm:mpch1} quantifies the spread of the eigenvalues around their mean $1$. Note that the range of the deviation increases when $\gamma$ increases from $0$ to $\infty$.

If $p/n\to 0$, the limiting spectral distribution in Theorem \ref{thm:mpch1} is the Dirac measure at $1$. After an appropriate transformation of the sample covariance matrix one can obtain the semi-circle law defined in \eqref{eq:semi-circle} as a non-degenerate limiting spectral distribution in this case.

The crucial assumptions in Theorem \ref{thm:mpch1} are the finiteness of the variance and that $p$ and $n$ tend to infinity at the same rate. By Theorem 2.8 in Bai \cite{Bai:1999}, the conclusion still holds if the entries are independent, have common mean and satisfy the Lindeberg-type condition
\begin{equation*}
\lim_{\nto} \frac{1}{\delta^2 n p}\sum_{i,t} \E[X_{it}^2 \1_{\{|X_{it}|>\delta \sqrt{n}  \}}] =0\,,\quad \delta>0\,.
\end{equation*}
\par

Many important test statistics in multivariate analysis are functions of the eigenvalues
$\la_{(1)}\ge \cdots \ge \la_{(p)}$ of the sample covariance matrix and can be expressed by means of the empirical spectral distribution of $n^{-1}\X\X'$.

\begin{example}{\em 
Let $(X_{it})$ be iid standard normal and consider
\begin{equation*}
V_n:=\log(\det (n^{-1}\X\X'))=
\log \prod_{i=1}^p \frac{\lambda_{(i)}}{n} =  p \int_{0}^{\infty} \log x \;\dint F_{n^{-1}\X\X'}(x)\,.
\end{equation*}
On the one hand, if $p$ is fixed we know from Example 1.1 in Yao et a.~\cite{yao:zheng:bai:2015} that 
\begin{equation}\label{eq:srgdrgd}
\sqrt{\frac{n}{p}}\, V_n \cid Y\sim N(0,2)\,,\quad \nto\,.
\end{equation}
On the other hand, if $p/n \to \gamma\in (0,1)$, one can use Theorem \ref{thm:mpch1} to obtain asymptotic values for $V_n$. On gets \as
\begin{equation*}
\frac{V_n}{p}\to \int_{0}^{\infty} \log x \;\dint F_{\gamma}(x)= \frac{\gamma-1}{\gamma} \log(1-\gamma)-1<0\,,
\end{equation*}
which implies $\sqrt{n/p}\, V_n \to -\infty$ \as~In view of \eqref{eq:srgdrgd} the asymptotic behavior of the test statistic $V_n$ crucially depends on the dimension $p$.}
\end{example}

A characterization of the limiting spectral distribution of sample covariance matrices with  general population covariance has been derived for many settings. As an example we state Theorem 1.1 in Bai and Zhou \cite{bai:zhou:2008}.
%a result that we will use in Chapter \ref{ch:corr} to find the limiting spectral distribution of sample correlation matrices.
\begin{theorem}\label{them:baizhou}
Assume \eqref{Cp} and the following conditions.
\begin{itemize}
\item For all $k$, $\E[X_{jk}X_{lk}]=T_{lj}$, and for any non-random $p\times p$ matrix $\bfB$ with bounded norm, $\E[(\x_k'\bfB \x_k -\tr(\bfB \bfT))^2]=o(n^{-2})$, where $\bfT=\bfT_n=(T_{jl})$ and $\x_k$ are the columns of $\X$.
\item The norm of $\bfT_n$ is uniformly bounded and $F_{\bfT_n}$ tends to a non-random probability distribution $H$.
\end{itemize}
Then, with probability $1$, $F_{n^{-1}\X\X'}$ tends to a probability distribution, whose Stieltjes transform $m(z)$ satisfies
\begin{equation}\label{eq:mpequation}
m(z)=\int_{\R} \frac{1}{t(1-\gamma-\gamma z m(z))-z} \dint H(t)\,.
\end{equation}
If $\underline{m}(z)=-(1-\gamma)/z+\gamma m(z)$, then \eqref{eq:mpequation} becomes
\begin{equation}\label{eq:silversteineq}
z=-\frac{1}{\underline{m}(z)}+\gamma \int_{\R} \frac{1}{1+\underline{m}(z) \,t} \dint H(t)\,.
\end{equation}
\end{theorem}
For historical reasons, \eqref{eq:mpequation} is often called {\em \MP equation}. In practice, the inversion of such integral equations can be very difficult. Therefore, a characterization of the limiting spectral distribution via \eqref{eq:mpequation} is of limited use. For numerical procedures, the {\em Silverstein equation} \eqref{eq:silversteineq} is sometimes preferred. 
Roughly speaking, the only known explicit examples of non-degenerate limiting spectral distributions are the \MP law, the circular law, the semi-circle law and the multivariate $F$-matrix. In contrast, the number of existence results is huge. 
We refer to the discussion in Yao et al.~\cite{yao:zheng:bai:2015} for further details.
\medskip

In Theorem \ref{thm:mpch1} we have seen that the sequence $(F_{n^{-1}\X \X'})$ converges to the Mar\v cenko--Pastur law if the iid entries possess a finite second moment. Now we will discuss the situation when the entries are still iid, but have an infinite variance. Here we assume the entries to be regularly varying with index $\alpha \in (0,2)$; see \eqref{eq:regvarch1} on page \pageref{eq:regvarch1} for the definition of regular variation.
Assuming \eqref{Cp} with $\gamma\in (0,1]$ in this infinite variance case, Belinschi et al.~\cite[Theorem~1.10]{belinschi:dembo:guionnet:2009} showed that the sequence $(F_{a_{n+p}^{-2}\X \X'})$ converges with probability one to a non-random probability measure with density $\rho_{\alpha}^\gamma$ satisfying
\begin{equation}\label{eq:esd}
\rho_{\alpha}^\gamma(x) x^{1+\alpha/2} \to \frac{\alpha \gamma}{2(1+\gamma)}, \quad x \to \infty\,;
\end{equation}
see also Ben Arous and Guionnet \cite[Theorem~1.6]{arous:guionnet:2008}.
Here the normalization $(a_k)$ is defined \sth\ 
\begin{equation}\label{eq:defanp}
\P(|X|>a_k)\sim k^{-1}\,,\quad k\to\infty\,.
\end{equation}
An application of the Potter bounds (see Bingham et al. \cite[p.~25]{bingham:goldie:teugels:1987}) shows that $a_{n+p}^2/n \to \infty$. To the best of our knowledge, explicit expressions or computational methods for the limiting spectral distribution in the infinite variance case are not available at this moment.

\section{Limits of extreme eigenvalues under finite fourth moment}

After the limiting spectral distribution for sample covariance matrices had been found, the focus shifted to the asymptotic behavior of the largest and smallest eigenvalues $\la_{(1)}$ and $\la_{(p)}$, respectively, of $\X\X'$. 
In this section, we assume that the entries of the data matrix $\X$ are iid with generic element $X$. Furthermore, suppose $p\le n$; otherwise $\la_{(p)}=0$ since $\X\X'$ has at most $\min(n,p)$ non-zero eigenvalues.
We will discuss the setting $\E[X^4]<\infty$, while the case $\E[X^4]=\infty$ is treated in Section \ref{sec:contr}.

Under condition \eqref{Cp} with $\gamma\le 1$, one can infer from Theorem \ref{thm:mpch1} that
\begin{equation}\label{eq:conlcu}
\liminf_{n\to \infty}\frac{\la_{(1)}}{n} \ge (1+\sqrt{\gamma})^2\quad \mbox{and}\quad
\limsup_{n\to \infty}\frac{\la_{(p)}}{n} \le (1-\sqrt{\gamma})^2\quad \as 
\end{equation}
We follow Bai and Silverstein \cite{bai:silverstein:2010} and derive necessary conditions for the \as~convergence of $n^{-1}\la_{(1)}$. Since the largest diagonal entry of a matrix is bounded by its largest eigenvalue, we have
\begin{equation}\label{eq:simplebo}
\frac{\la_{(1)}}{n}\ge \max_{i=1,\ldots,p} \frac{1}{n} \sum_{t=1}^n X_{it}^2\,.
\end{equation}
If $\E[X^4]=\infty$, then by Lemma B.25 in \cite{bai:silverstein:2010}
\begin{equation}\label{eq:sghdfgdfs}
\limsup_{n\to \infty}\max_{i=1,\ldots,p} \frac{1}{n} \sum_{t=1}^n X_{it}^2 =\infty \quad \as
\end{equation}
%Next note that Theorem \ref{thm:mpch1} holds if $\E[X]=c\neq 0$. 
If $\E[X^4]<\infty$ and $\E[X]=c\neq 0$, then
\begin{equation*}
\frac{1}{\sqrt{n}}\twonorm{\X}\ge \frac{1}{\sqrt{n}}\twonorm{\E[\X]}  -\frac{1}{\sqrt{n}}\twonorm{\X-\E[\X]}
\ge \frac{|c| p}{\sqrt{n}}  - \frac{1}{\sqrt{n}}\twonorm{\X-\E[\X]} \to\infty \quad \as,
\end{equation*}
where for any matrix $\bfA$, $\twonorm{\bfA}$ denotes its spectral norm, i.e., its largest singular value. This and \eqref{eq:sghdfgdfs} show that $\E[X]=0$ and $\E[X^4]<\infty$ are necessary conditions for the \as~convergence of $n^{-1}\la_{(1)}$.

{\em In what follows, we assume $\E[X]=0$ and $\E[X^2]=1$, whenever the respective moments exist.}

\subsection{Sample covariance matrices}
Extending work by Geman \cite{geman},
Bai et al.~\cite{bai:yin:krishnaiah:1988} 
showed under \eqref{Cp} that
\beam\label{eq:gemanch1}
\dfrac {\la_{(1)}}{n} \stas \big(1+\sqrt{\gamma}\big)^2\,,\qquad \nto\,,
\eeam
which is the optimal result in view of \eqref{eq:sghdfgdfs}. Later Bai and Yin \cite{bai:yin:1993} proved the following result under the additional assumption $\gamma\in (0,1)$:
\beam\label{eq:jo45}
\limsup_{n\to \infty} \twonorm{n^{-1} \X\X'-(1+\gamma) \bfI} \le 2 \sqrt{\gamma}\,\quad \as
\eeam
Because of 
\begin{equation*}
\twonorm{n^{-1} \X\X'-(1+\gamma) \bfI} = \max \{ \la_{(1)}/n- (1+\gamma), -\la_{(p)}/n+(1+\gamma)\}\,,
\end{equation*}
equations \eqref{eq:jo45} and \eqref{eq:conlcu} imply 
\begin{equation}\label{eq:sdfgsdfdf}
\lim_{n\to \infty}\frac{\la_{(1)}}{n} = (1+\sqrt{\gamma})^2\quad \mbox{and}\quad
\lim_{n\to \infty}\frac{\la_{(p)}}{n} = (1-\sqrt{\gamma})^2\quad \as \,
\end{equation}
The approach based on \eqref{eq:jo45} provides a lower bound on the smallest eigenvalue. Unfortunately, one cannot gain any information about the minimal conditions for the existence of a limit of $\la_{(p)}/n$ since the method treats $\la_{(1)}$ and $\la_{(p)}$ simultaneously. Therefore it can (at best) only be applied under $\E[X^4]<\infty$. Consequently, one loses sharpness of the conditions for the limit of $\la_{(p)}/n$. It was finally discovered by Tikhomirov \cite{tikhomirov:2015} that the \as~limit of $ \la_{(p)}/n$ is given by \eqref{eq:sdfgsdfdf} if $\E[X^2]=1$, whereas higher moments can be infinite.
\par

Under suitable moment assumptions, $\la_{(1)}$ and $\la_{(p)}$ possess {\em Tracy--Widom} fluctuations around their almost sure limits. For instance, 
Johnstone \cite{johnstone:2001} complemented \eqref{eq:sdfgsdfdf} by the corresponding \clt\ in the special case of iid standard normal 
entries:
\begin{equation*}%\label{eq:john}
 n^{2/3}\,\dfrac{(\sqrt{\gamma})^{1/3}}{\big(1+\sqrt{\gamma}\big)^{4/3}}\Big(\dfrac {\la_{(1)}}{n} -
\big(1+\sqrt{\tfrac pn }\big)^2\Big)
\std \xi\,,
\end{equation*}
where the limiting \rv\ has a {\em Tracy--Widom \ds} of order~$1$.
Its distribution function $G_1$ is given by
\begin{equation*}
G_1(s) = \exp\Big\{
  -\frac{1}{2} \int_{s}^\infty [
    q(x) + (x - s) q^2(x)
 ] \dint x
\Big\}\,,
\end{equation*}
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{equation*}
  q''(x) = xq(x) + 2 q^3(x)\,,
\end{equation*}
where $ q(x)\sim {\rm Ai}(x)$ as $x \to \infty$ and Ai$(\cdot)$ is the Airy kernel; see Tracy and Widom~\cite{tracy:widom:2012} for details.

\subsection{Sample correlation matrices}

In comparison with the eigenvalues of $\X\X'$, much less is known about the ordered eigenvalues
\begin{equation*}
\mu_{(1)} \ge \cdots \ge\mu_{(p)}\,
\end{equation*} 
of the {\em sample correlation matrix} $\bfR=\Y\Y'$ with entries
\begin{equation}\label{eq:corrRch1}
R_{ij}=\sum_{t=1}^n \frac{X_{it}X_{jt}}{\sqrt{D_i} \sqrt{D_j}} = \sum_{t=1}^n Y_{it}Y_{jt}\,, \quad i,j=1,\ldots,p\,.
\end{equation}
In this thesis we will often make use of the notation $\Y=(Y_{it})=(X_{it}/\sqrt{D_i})$ and
\begin{equation}\label{eq:Dch1}
D_i=D_i^{(n)}=\sum_{t=1}^n X_{it}^2\,, \qquad
i=1,\ldots,p;\;  n\ge 1\,.
\end{equation}

With $\bfF= \diag (1/D_1,\ldots,1/D_p)$, we have $\bfR={\bfF}^{1/2} {\X}{\X}'{\bfF}^{1/2}$ which has the same eigenvalues as ${\X}{\X}'{\bfF}$. Weyl's inequality (see \cite{bhatia:1997}) yields
\begin{equation}\label{eq:lamuch1}
\begin{split}
\max_{i=1,\ldots,p} |\mu_{(i)}- n^{-1} \la_{(i)}| &\le \twonorm{{\X}{\X}'{\bfF}-n^{-1}{\X}{\X}' }\\
&\le n^{-1} \twonorm{{\X}{\X}'}  \twonorm{n \bfF- \bfI}\\
&= n^{-1} \la_{(1)} \max_{i=1,\ldots,p} \Big| \frac{n}{D_i} -1 \Big|\,,
\end{split}
\end{equation}
which converges \as~to $0$ if $\E[X^4]<\infty$; see Chapter~\ref{ch:corr} for details.
Under \eqref{Cp} with $\gamma\in(0,1]$,  Jiang \cite{jiang:2004}, and Xiao and Zhou \cite{xiao:zhou:2010} used this approach to derive 
\begin{equation*}
\mu_{(1)} \to (1+\sqrt{\gamma})^2\, \quad \text{ and } \quad  \mu_{(p)} \to (1-\sqrt{\gamma})^2\, \quad \as
\end{equation*}

\section{Eigenvectors}

Eigenvectors of large random matrices and graphs play an essential role in statistical analysis, physics and computer science. Many properties of a matrix or a graph can be deduced from its eigenvectors. Popular algorithms for data analysis such as spectral clustering, principal component analysis, PageRank and community detection are based on the eigenvector-eigenvalue decomposition of a matrix.

If the data matrix $\X$ has iid standard normal entries, then $\X\X'$ is a {\em Wishart matrix}, whose eigenvectors are well studied; see for example Bai and Silverstein \cite[Ch.~10]{bai:silverstein:2010}. Due to the invariance of $\X\X'$ under orthogonal transformations, its matrix of properly normalized eigenvectors is Haar distributed, i.e., the distribution is uniform on the space of orthogonal $p\times p$ matrices. This result has been extended to different classes of matrices $\X$ by direct comparison with Wishart matrices. Such statements are called universality results and they often require that the new entry distribution is in some sense similar to the standard normal distribution. This is often achieved by moment conditions. Silverstin \cite{silverstein:1984} showed that the matrix of eigenvectors is asymptotically Haar distributed as $p/n\to \gamma\in(0,\infty)$ if the first four moments of the iid entries coincide with those of the standard normal distribution. On the one hand, this means that eigenvectors of $\X\X'$ are completely unstructured for light-tailed entry distributions.  On the other hand, the extreme eigenvalues converge to constants \as 
\par

Although eigenvectors play a minor role in this thesis, we summarize some results from the literature on the light-tailed case.  Consequently, our approximations of eigenvectors in the heavy-tailed case are put into context. The majority of studies on eigenvectors of large random matrices is conducted on Wigner matrices $\bfW=(W_{ij})$. They are symmetric, real-valued, $n\times n$ matrices with entries $W_{ij}, 1\le i \le j \le n$, that are iid, mean zero and unit variance random variables.

Roughly speaking, the spectral properties of the square of a Wigner matrix and a sample covariance matrix with $p=n$ are quite similar if they share the same sufficiently light-tailed entry distribution. For the purpose of exposition, we look at $\bfW^2=\bfW \bfW'$ instead of $\X\X'$. Indeed, apart from the additional symmetry restriction for $\bfW$ they have the same structure.

By Theorem 2.1 in \cite{bai:fang:liang:2014}, the limiting spectral distribution of $(\bfW /\sqrt{n})$ is the {\em semi-circle law} $G$ with density
\begin{equation}\label{eq:semi-circle}
g(x)=\frac{1}{2\pi} \sqrt{4-x^2}\, \1_{\{ |x|\le 2 \}}\,.
\end{equation}
The semi-circle law and the \MP law $F_{\gamma}$ with $\gamma=1$ are linked in the following way: if $Y\sim G$ then $Y^2\sim F_1$.

Since $\bfW^2$ and $\bfW$ have the same eigenvectors it will be sufficient to study the latter. In the remainder of this section we list some properties of $\bfv_1, \ldots, \bfv_n$, where $\bfv_i$ is an eigenvector associated to the $i$th largest eigenvalue of $\bfW$. In addition, we assume $\bfv_i$ are unit vectors, i.e. $\ltwonorm{\bfv_i}=1$, and that the first non-zero component of $\bfv_i$ is positive.

Often the entries are assumed to be {\em sub-exponential}. We call a random variable $W$ {\em sub-exponential} with exponent $\alpha>0$ if there exists a constant $\beta>0$ such that 
\begin{equation*}
\P(|W|>x)\le \beta \exp(-x^\alpha /\beta)\,,\quad x>0\,.
\end{equation*}

Theorem 1.4 in O'Rourke et al.~\cite{orourke:vu:wang:2016} focuses on Wigner matrices with sub-exponential entries with exponent $\alpha$. There exists a constant $C_{\alpha}>0$ such that the probability that the spectrum of $\bfW$ is simple and that every coordinate of every $\bfv_i$ is non-zero is at least $1-C_{\alpha} n^{-\alpha}$. By our sign convention, the eigenvectors are unique with high probability. In particular, if $W_{11}$ is standard normal, $\bfv_i$ is uniformly distributed on 
\begin{equation*}
\mathcal{S}_+^{n-1}:=\{ \x=(x_1,\ldots,x_n)': \ltwonorm{\x}=1 \text{ and } x_1>0 \}\,.
\end{equation*}
An eigenvector of a Wigner matrix with light-tailed entry distribution (for instance sub-exponential) behaves like a random vector uniformly distributed on $\mathcal{S}_+^{n-1}$. More precise quantitative statements are difficult to obtain. For details we refer to \cite{orourke:vu:wang:2016}.

If $\bfv=(v_1,\ldots,v_n)'$ is a random vector uniformly distributed on $\mathcal{S}_+^{n-1}$, probabilistic bounds on its coordinates are available. By Theorem 2.1 in \cite{orourke:vu:wang:2016}, we have 
\begin{equation}\label{eq:coordinates}
v_{\text{max}} := \max_{i=1,\ldots,n} |v_i| \le C\sqrt{\frac{\log n}{n}}\quad \text{ and } \quad v_{\text{min}} := \min_{i=1,\ldots,n} |v_i|\ge \frac{c}{n^{3/2}}
\end{equation}
with probability $1-o(1)$ for any $C>1$ and $c\in [0,1)$.

In RMT it is common to study the so-called {\em bulk} and {\em edge} spectra separately.
For $\vep\in (0,1)$ one distinguishes between the {\em bulk} eigenvectors $\bfv_i, i\in \{1\le t\le n: \vep n \le t \le (1-\vep) n\} :=B_\vep$ and the {\em edge} eigenvectors $\bfv_i, i \in \{1,\ldots,n\} \backslash B_\vep$. The associated eigenvalues are usually referred to as bulk and edge spectrum, respectively. Roughly speaking, the limiting spectral distribution of a sequence of random matrices depends on the bulk spectrum, while the edge spectrum influences the behavior of functionals of the eigenvalues.
\par

The behavior of the largest coordinates of $\bfv_1,\ldots, \bfv_n$ was studied in the case of certain sub-exponential entries with exponent $\alpha=2$ whose first four moments match those of the standard normal distribution. 
By Theorems 4.1 and 4.3 in \cite{orourke:vu:wang:2016}, one has for any bulk eigenvector $\bfv_i$ and appropriate constants $c_1,c_2>0$,
\begin{equation*}
c_1 \sqrt{\frac{\log n}{n}} \le \bfv_{i,\text{max}}\le c_2 \sqrt{\frac{\log n}{n}}
\end{equation*}
with probability $1-o(1)$. This result is astonishingly precise. Note that up to logarithmic corrections $\bfv_{i,\text{max}}$ is of the smallest possible magnitude $n^{-1/2}$. This property is called {\em complete delocalization}; see \cite{rudelson:vershynin:2015}.
For edge eigenvectors, however, it is proved that for a constant $c_3>0$,
\begin{equation*}
\bfv_{i,\text{max}}\le c_3 \frac{\log n}{\sqrt{n}}\,
\end{equation*}
with probability $1-o(1)$.
The optimal bound in the edge case remains an open problem. Additionally, by Corollary 5.4 in \cite{orourke:vu:wang:2016} the $\ell_p$-norms, $1\le p\le2$, of the $(\bfv_i)$ are of the same order of magnitude:
\begin{equation}\label{eq:lpefgs}
c_0 n^{1/p-1/2}\le \min_{i=1,\ldots,n} \| \bfv_i\|_{\ell_p} \le \max_{i=1,\ldots,n} \| \bfv_i\|_{\ell_p}  \le C_0 n^{1/p-1/2}
\end{equation}
with probability $1-o(1)$ for positive constants $c_0,C_0$.
\par

If the entry distribution of $\bfW$ has heavy tails, the behavior of its eigenvectors is completely different. In \cite{benaych:peche}, Benaych-Georges and P\'{e}ch\'{e} assumed $W_{11}$ to be regularly varying with index $\alpha\in (0,4)$; see \eqref{eq:regvarch1} on page \pageref{eq:regvarch1}, and provided an approximation of $\bfv_k$ for any fixed $k$. From their asymptotic result one can deduce that $\bfv_{k,\text{max}}$ converges to $1/\sqrt{2}$ in probability. Moreover, asymptotically there are only two coordinates of $\bfv_k$ with non-zero mass. Both are of magnitude $1/\sqrt{2}$. This is the opposite of complete delocalization: {\em complete localization}. The number of non-zero coordinates is bounded.

Note that in the presence of heavy tails the eigenvectors of $\bfW$ and $\X\X'$ are very different; see \eqref{eq:eigenvec} in Section \ref{sec:contr}.

>>>>

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: If a function $f(\cdot)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\]
then we say $f(x)$ is regularly varying with index $\alpha$.
Distribution functions, say $F(\cdot)$, whose survival function
$\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
observed in a variety of phenomena, for example, phase transition of
matter, black body radiation, neuronal avalanches and stock returns.
When expanded to multiple dimensions, the scaling property is better
described in terms of weak convergence to a spectral measure: If a
random vector $V$ satisfies
\[
\lim_{x \to \infty}
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
}
\overset{w}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\]
then we say the components of $V$ is jointly regularly varying with
index $\alpha$. Here $\mu_\alpha(\cdot)$ is a probability measure on
the unit sphere \cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is jointly regularly varying with index $\alpha$, then each component
of it and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et al
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

Thus estimating the tail index $\alpha$ is particularly important for
understanding the behaviour of a heavy-tailed series. The earliest
method proposed for this purpose is due to Hill \cite{hill1975simple}:
\[
\hat \alpha_H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
  X_{(i)} \over X_{(k+1)}
  \right)
  \right]^{-1}
\]
where $X_1, X_2, \dots$ is the series whose tail index is the subject
of estimation, and $X_{(i)}$ is the $i$-th upper order statistics of
this series. Several authors have contributed to showing the
weak consistency of the estimator $\hat \alpha_H$,
i.e. if $k \to \infty, k/n \to 0$ as $n \to \infty$, then
$\hat \alpha_H \overset{P}{\to} \alpha$.

Mason \cite{mason:1982} first proved that the estimator
was consistent when $X_1, X_2, \dots$ were iid; later Rootz\'en,
Leadbetter and de Haan \cite{rootzen:leadbetter:dehaan1992} and Hsing
\cite{hsing:1991} proved its consistency when $X_1, X_2, \dots$ were
weakly dependent; and Resnick and St{\u a}ric{\u a}
\cite{resnick:starica:1995, resnick:starica:1997} proved its
consistency when $X_t$ was a linear process.
